### 2019-08-13 ops/tex/abstract.tex
Update

\begin{abstract}
In distributed computing frameworks like MapReduce and Spark, shuffle phases are barriers between the adjacent computing phases. 
Since the transmission of the massive data, shuffle phases account for more than 33\% of job completion time.
In this paper, we optimize shuffle phases in two ways: 
(1) we decouple the shuffle from the computing phases to increase the resource utilization, and
(2) we optimize the network scheduling to decrease the communication time in the shuffle phases.
We present \textit{OPS}, a system that enables the common computing frameworks to use scheduling algorithms to optimize the whole data shuffle progress.
We show that \textit{OPS} improves the end-to-end job completion time by up to ??.
We also show \dots
\end{abstract}